---
title: "Create Seurat Object for Menon et al 2020 dataset"
author: "Javier Perales-Patón - javier.perales@bioquant.uni-heidelberg.de - ORCID:0000-0003-0780-6683"
output: github_document
---

## Setting-up environment
The environment will be set with a random seed number for reproducibility and 
an output folder for processed data and figures.

### set env
```{r env}
options(stringsAsFactors = FALSE)
# Seed number
set.seed(1234)
# Output directory
OUTDIR <- "./01_Seurat/"
if(!dir.exists(OUTDIR)) dir.create(OUTDIR);

# Figures
FIGDIR <- paste0(OUTDIR, "/figures/")
knitr::opts_chunk$set(fig.path=FIGDIR)
knitr::opts_chunk$set(dev=c('png','tiff'))
# Data
DATADIR <- paste0(OUTDIR, "/data/")
if(!dir.exists(DATADIR)) dir.create(DATADIR);
```

### Load libraries
Essential libraries for R analysis.
```{r}
library("dplyr")
library("Seurat")
library("ggplot2")
```

## Load data
Individual samples are stored in individual folders in the `./raw/scRNAseq`.

We first load the metadata of all samples.
```{r meta}
meta <- read.table("./data/scRNAseq_metadata.tsv", sep="\t", header=TRUE, stringsAsFactors=FALSE, quote = "")

# We create a dictionary of sampleID from raw to supplementary files
sID_dict <- setNames(meta$sampleID, meta$sampleID_suppl)
print(sID_dict)
```
And the metadata for individual cells

```{r}
meta2 <- read.table("data/PREMIERE_TIS_JUNE2019_CELLBARCODES_CLUSTER.txt", sep="\t", header=TRUE,
		    stringsAsFactors = FALSE)
# We recover the original SampleID
for(sID_suppl in names(sID_dict)) {
	sID_actual <- sID_dict[sID_suppl]
	meta2$cell <- gsub(paste0("^",sID_suppl,"_"),
			    paste0(sID_actual,"_"),
			    meta2$cell)
}
cellOrigin <- sapply(meta2$cell, function(x) strsplit(x, split="_")[[1]][1])

# Fix one problematic sample ID
cellOrigin <- gsub("^KPMP", "KPMP_Pilot_18342", cellOrigin)

cellcluster_bySample <- split(setNames(meta2$cluster.name.abbreviation, meta2$cell),
				       cellOrigin)
```


```{r, fig.width=14, fig.height=6}
dirs <- list.dirs("./raw/scRNAseq/")
# We remove the first element that is the folder itself
dirs <- dirs[-1]
# Each folder is an individual sample
names(dirs) <- basename(dirs)

SS <- setNames(vector("list", length=length(dirs)), names(dirs))
for(GSM in names(SS)) {
	cat(paste0("Processing ",GSM, "\n"), file=stdout())
	idx <- which(meta$geo_accession==GSM)
	sname <- meta[idx, "sampleID"]
	sname2 <- meta[idx, "sampleID_suppl"]
	biopsy <- meta[idx, "biopsy"]
	mtcutoff <- meta[idx, "processing.mtcutoff"]
	nFeature.min <- meta[idx, "processing.nFeature.min"]
	nFeature.max <- meta[idx, "processing.nFeature.max"]
	chemistry10X <- meta[idx, "chemistry10X"]

# 	tag <- meta[idx, "tag"]


	fl <- dirs[GSM] 
	names(fl) <- sname

	dat <- Read10X(data.dir=fl)
	S <- CreateSeuratObject(counts = dat, project = sname, 
				min.cells=3, min.features=200)
	S[["percent.mt"]] <- PercentageFeatureSet(S, pattern = "^MT-")

	## QC metrics
	# Visualize QC metrics as a violin plot
	VlnPlot(S, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

	# Feature scatter
	plot1 <- FeatureScatter(S, feature1 = "nCount_RNA", feature2 = "percent.mt") + 
		geom_hline(yintercept=c(0, mtcutoff))
	plot2 <- FeatureScatter(S, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") +
		geom_hline(yintercept=c(nFeature.min, nFeature.max))
	print(plot1 + plot2)

	# Filter
	S <- subset(S, subset= nFeature_RNA > nFeature.min & nFeature_RNA < nFeature.max &
		    percent.mt < mtcutoff)

	# run sctransform
# 	Note that this single command replaces NormalizeData, ScaleData, and FindVariableFeatures.
# 	Transformed data will be available in the SCT assay, which is set as the default after running sctransform
# 	During normalization, we can also remove confounding sources of variation, for example, mitochondrial mapping percentage
	S <- SCTransform(S, vars.to.regress = "percent.mt", verbose = FALSE)

	# Add some metadata
	S$chemistry10X <- chemistry10X
	S$biopsy <- biopsy

	# Add cluster from original publication
	cidx <- gsub("-1$","",colnames(S))
	cl <- setNames(cellcluster_bySample[[sname]][cidx], 
		       colnames(S))
	S$Abb <- cl
	# GEOACC
	S$GSMACC <- GSM

	# Save the data
	saveRDS(S, file=paste0(DATADIR,"/",GSM,".rds"))

}
```

## Session Info
```{r}
sessionInfo()
```

