---
title: "Create Seurat Object"
author: "Javier Perales-Patón - javier.perales@bioquant.uni-heidelberg.de - ORCID:0000-0003-0780-6683"
output: 
  github_document:
    html_preview: false
params:
  ID: 
    value: x
  OUTDIR:
    value: x
---

```{r, include=FALSE}
# params <- list(ID="GSM4191961", OUTDIR="./01_Seurat")
```

## Setting-up environment
The environment will be set with a random seed number for reproducibility and 
an output folder for processed data and figures.

### set env
```{r env}
options(stringsAsFactors = FALSE)
# Seed number
set.seed(1234)
# Output directory
OUTDIR <- params$OUTDIR
if(!dir.exists(OUTDIR)) dir.create(OUTDIR);

# Figures
FIGDIR <- paste0(OUTDIR, "/figures/")
knitr::opts_chunk$set(fig.path=FIGDIR)
knitr::opts_chunk$set(dev=c('png'))
# Data
DATADIR <- paste0(OUTDIR, "/data/")
if(!dir.exists(DATADIR)) dir.create(DATADIR);
```

### Load libraries
Essential libraries for R analysis.
```{r}
library("dplyr")
library("Seurat")
library("ggplot2")
library("yaml")
library("DoubletFinder")
source("../src/seurat_fx.R")
```

## Load data
Individual samples are stored in individual folders in the `./00_scRNAseq`.


```{r read_idx}
# Find index file
yaml_fl <- paste0("../index/samples/",params$ID,".yaml")
idx <- yaml::read_yaml(yaml_fl)
# Reformat for easy access to info
idx$seurat_cutoff <- unlist(lapply(idx$seurat_cutoff, function(z) setNames(z[[1]], names(z))))
idx$metadata <- unlist(lapply(idx$metadata, function(z) setNames(z[[1]], names(z))))
```

```{r}
cat(paste0("Processing ",params$ID, "\n"), file=stdout())
# Find folder of 10x cellranger count
input_dir <- paste0("./00_scRNAseq/",params$ID)
names(input_dir) <- params$ID

dat <- Read10X(data.dir=input_dir)
S <- CreateSeuratObject(counts = dat, project = params$ID, 
			min.cells=5, 
			min.features=idx$seurat_cutoff["MIN_nFeatures"])
S[["percent.mt"]] <- PercentageFeatureSet(S, pattern = "^MT-")
```

## QC
```{r QC, fig.width=14, fig.height=6}
# Visualize QC metrics as a violin plot
VlnPlot(S, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

# Feature scatter
plot1 <- FeatureScatter(S, feature1 = "nCount_RNA", feature2 = "percent.mt") + 
		geom_hline(yintercept=c(0, idx$seurat_cutoff["MAX_percMT"]))
plot2 <- FeatureScatter(S, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") +
		geom_hline(yintercept=c(idx$seurat_cutoff["MIN_nFeatures"], 
					idx$seurat_cutoff["MAX_nFeatures"]))
print(plot1 + plot2)
```

## Filter-out
```{r}
# Before
before_nCells <- ncol(S)
# apply Filter
S <- subset(S, subset= nFeature_RNA > idx$seurat_cutoff["MIN_nFeatures"] & 
	    		nFeature_RNA < idx$seurat_cutoff["MAX_nFeatures"] &
	    percent.mt < idx$seurat_cutoff["MAX_percMT"])
# After
after_nCells <- ncol(S)

print(c(before_nCells, after_nCells))
```

## Normalization using scTransform
```{r warning=FALSE}
#NOTE: this single command replaces NormalizeData, ScaleData, and FindVariableFeatures.
# Transformed data will be available in the SCT assay, which is set as the default after running sctransform
# During normalization, we can also remove confounding sources of variation, for example, 
# mitochondrial mapping percentage
S <- SCTransform(S, vars.to.regress = "percent.mt", verbose = FALSE)
```

## Add metadata
```{r}
for(feat in names(idx$metadata)) {
	    S[[feat]] <- idx$metadata[[feat]]
}
```

## Finding Variable genes
```{r HGV}
S <- FindVariableFeatures(S, selection.method = "vst", nfeatures = 2000)
```

```{r PCA}
S <- RunPCA(S, npcs=50, features=VariableFeatures(S))
calcNPCs <- get_npcs(S)
(nPCs <- calcNPCs$npcs)
ElbowPlot(S) + geom_vline(xintercept = nPCs)
```

```{r UMAP}
S <- RunUMAP(S, dims= 1:nPCs)
```

## Doublet detection (using DoubletFinder)

```{r pK_id, message=FALSE, warning=FALSE}
# No ground-truth: pK identification
sweep.res.list <- paramSweep_v3(S, PCs = 1:nPCs, sct = TRUE)
sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)
bcmvn <- find.pK(sweep.stats)
# Do I have to select here pK based on something??
# > Optimal pK for any scRNA-seq data can be manually discerned as maxima in BCmvn distributions
# NOTE: pK is encoded as a factor. But actually is a number. If you transform
# a factor level to a number directly you will get level position. So as.char is needed.
pK_sel <- as.numeric(as.character(bcmvn[which.max(bcmvn$BCmetric), "pK"]))
```

```{r doublet_prop_est}
## Assuming 0.75% doublet formation rate per each 1000 cells in 10x run.
## src: https://assets.ctfassets.net/an68im79xiti/1eX2FPdpeCgnCJtw4fj9Hx/7cb84edaa9eca04b607f9193162994de/CG000204_ChromiumNextGEMSingleCell3_v3.1_Rev_D.pdf
nExp_poi <- (before_nCells/1e3 * 0.0075)  
S <- doubletFinder_v3(S, PCs = 1:nPCs, pN = 0.25, 
			       pK = pK_sel, 
			       nExp = nExp_poi, reuse.pANN = FALSE, 
			       sct = TRUE)
# Rename cols
DF_idx <- grep("^(pANN_|DF\\.)", colnames(S@meta.data))
colnames(S@meta.data)[DF_idx] <- gsub("^(pANN|DF)(_|\\.)?.*","\\1",colnames(S@meta.data)[DF_idx])
```

How many doublets were found?
```{r doublets, message=FALSE, warning=FALSE}
table(S@meta.data$DF) # DF stands for DoubletFinder
```

## Save object
```{r}
saveRDS(S, paste0(DATADIR,"/S.rds"))
```

## sessionInfo
```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
